\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-05-21 doc/paper.Rmd
starting proofing

## Motivation
Same source analyses are a major part of an Forensic Toolmark Examiner's job. In current practice examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}. Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

@@

\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
\gk{table included}
Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.

\hh{We need an additional sentence here to get from the data to the statistical methods ... }
\gk{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}
\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach is non-deterministic, because segments are chosen randomly.  \citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically. This approach also ensures independence between segements of striae.  In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
\end{table}

>>>>

## Motivation
Same source analyses are a major part of a Forensic Toolmark Examiner's job. In current practice, examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}. Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

@@

Various toolmarks have been studied in the literature (see \autoref{tab:toolmarks-ER} for an overview):  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.

Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as basis for an error rate calcluation.
\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach is non-deterministic, because segments are chosen randomly.  \citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically. This approach also ensures independence between segements of striae.  In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. All reported papers use some variation of the Chumbley-score method.}
\end{table}

### 2018-05-21 doc/paper.tex
starting proofing

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

@@

Same source analyses are a major part of an Forensic Toolmark Examiner's
job. In current practice examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
These conclusions are made on the basis of ``unique surface contours''
of the two toolmarks being in ``sufficient agreement''
\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
agreement'' as the possibility of another tool producing the markings
under comparison, as practically impossible \citep{afte-toolmarks1998}.
Potential subject bias in the assessment as well as the lack of
specified error rates are the main points of criticisms first raised by
the National Research Council in 2009 \citep{NAS:2009} and later
emphasized further by the President's Council of Advisors on Science and
Technology \citep{pcast2016}.

@@

{\textcolor{orange}{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}}
{\textcolor{blue}{table included}}

Various toolmarks have been studied in the literature:
\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
have investigated 3D marks from screwdriver, tongue and groove pliers
captured using a confocal microscope; \citet{afte-chumbley} have been
investigated digitized marks from slip-joint pliers generated by a
surface profilometer.

{\textcolor{orange}{We need an additional sentence here to get from the data to the statistical methods ... }}
{\textcolor{blue}{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}}

\citet{manytoolmarks2} define a relative distance metric and use it as
similarity measure between two toolmarks. \citet{manytoolmarks1} extract
many small segments in the markings of two toolmarks and compare
similiarity using a maximum pearson correlation coefficient. The
Chumbley scoring method, first introduced by \citet{chumbley}, uses a
similar but more extensive framework based on a Mann-Whitney U test of
the resulting correlation coefficients. This approach is
non-deterministic, because segments are chosen randomly. \citep{hadler}
make the score deterministic for each pair of toolmarks by choosing
segments for comparison systematically. This approach also ensures
independence between segements of striae. In this paper, we are
investigating the applicability of the Chumbley scoring method by
\citet{hadler} to assess striation marks on bullet lands for same-source
identification.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
\end{table}

@@

{\textcolor{blue}{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests. Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests. A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important (sub-)class characteristics}}

>>>>

\noindent%
{\it Keywords:} forensic science, toolmark, cross-correlation, Mann-Whitney U statistic, , 
\vfill

@@

Same source analyses are a major part of a Forensic Toolmark Examiner's
job. In current practice, examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
These conclusions are made on the basis of ``unique surface contours''
of the two toolmarks being in ``sufficient agreement''
\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
agreement'' as the possibility of another tool producing the markings
under comparison, as practically impossible \citep{afte-toolmarks1998}.
Potential subject bias in the assessment as well as the lack of
specified error rates are the main points of criticisms first raised by
the National Research Council in 2009 \citep{NAS:2009} and later
emphasized further by the President's Council of Advisors on Science and
Technology \citep{pcast2016}.

@@

Various toolmarks have been studied in the literature (see
\autoref{tab:toolmarks-ER} for an overview): \citet{manytoolmarks1} and
\citet{chumbley} have been analyzing screwdriver marks digitized using a
profilometer; \citet{manytoolmarks2} have investigated 3D marks from
screwdriver, tongue and groove pliers captured using a confocal
microscope; \citet{afte-chumbley} have been investigated digitized marks
from slip-joint pliers generated by a surface profilometer.

Analysis of these digitized markings require the use of statistical
methods which can quantify the scientific mechanism of comparing
markings and serve as basis for an error rate calcluation.
\citet{manytoolmarks2} define a relative distance metric and use it as
similarity measure between two toolmarks. \citet{manytoolmarks1} extract
many small segments in the markings of two toolmarks and compare
similiarity using a maximum pearson correlation coefficient. The
Chumbley scoring method, first introduced by \citet{chumbley}, uses a
similar but more extensive framework based on a Mann-Whitney U test of
the resulting correlation coefficients. This approach is
non-deterministic, because segments are chosen randomly. \citep{hadler}
make the score deterministic for each pair of toolmarks by choosing
segments for comparison systematically. This approach also ensures
independence between segements of striae. In this paper, we are
investigating the applicability of the Chumbley scoring method by
\citet{hadler} to assess striation marks on bullet lands for same-source
identification.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates in same-source toolmark analysis reported in the literature. All reported papers use some variation of the Chumbley-score method.}
\end{table}

@@

{\textcolor{blue}{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests. Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests. A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important individual characteristics}}
